Repeat Yourself? When Linux kernel maintainers balance Code Duplication and Readability

### 2025-10-20 paper/techdebt2026/techdebt.tex
techdebt2025: reorganizes methods and results
Signed-off-by: Paulo Meirelles <paulormm@ime.usp.br>

\section{Anonymous-tool}

We developed \texttt{anonymous-tool}, a command-line tool designed to detect and analyze function-level code duplication within the Linux kernel to support our investigation. Released under the MIT license, \texttt{anonymous-tool} employs a two-stage architecture that separates the computationally intensive task of finding duplicates from the process of querying the results. The first stage, the Preprocessor, analyzes the entire codebase, parses C functions, and stores identified duplicates in a database. The second stage, the Query Responder, uses this database to retrieve information about the identified duplications quickly.
%The tool is available at \textit{\href{https://github.com/arkanjo-tool/arkanjo}{github.com/arkanjo-tool/arkanjo}}.

\texttt{Anonymous-tool} is designed for the specific challenges of the kernel, with a focus on scalability to handle its size and prioritizing duplications that impact maintainability. The tool's primary role in this study was to provide a concrete entry point for new contributors by identifying function-level duplications to address. Figure~\ref{fig:diagrama} illustrates the tool's architecture. The tool uses a text similarity method that applies TF-IDF vector embedding, implemented with the Gensim library~\cite{gensim}, and computes cosine similarity to detect clones.

\begin{figure}[ht]
\centering
\includegraphics[scale=0.275]{diagrama_mestrado}
\caption{Architecture diagram with the tool components
}
\label{fig:diagrama}
\end{figure}

\section{Participant Observation Study}
\label{sec:participant}

@@

The first duplicated function pair chosen is function \textit{offset\_to\_id} on files \textit{dc/gpio/dcn32/hw\_translate\_dcn32.c} and \textit{dc/gpio/dcn315/hw\_translate\_dcn315.c}. The second duplicated function pair is function \textit{phy\_id\_to\_atom} in file 
\textit{ \seqsplit{dc/bios/dce110/command\_table\_helper\_dce110.c}} and \textit{\seqsplit{dc/bios/dce60/command\_table\_helper\_dce60.c}}.
Hereafter, the mitigation on \textit{offset\_to\_id} will be referred to as mitigation 1, and the mitigation on 
\textit{phy\_id\_to\_atom} will be referred to as mitigation 2.

@@

\subsection{Results}
\label{subsec:participantresults}

\subsubsection{Mitigation 1}

In the systematic approach, we found that the \textit{offset\_to\_id} function 
exists in 10 code files, enclosing multiple GPU architectures. The duplicated 
functions are not exactly equal, as some minimal specific logic is applied to each GPU architecture. To address this issue, we used the extract method to 
break the function into smaller parts and the parameterized method to address 
the specific logic for the GPUs.

\subsubsection{Mitigation 2}

Using the systematic approach, we found that the \textit{phy\_id\_to\_atom} function exists in five code files, and two other functions are duplicated 
in those five code files. All functions across the files were exactly equal, 
so there was no need to apply the refactoring methods presented in the literature. 
The refactoring was resumed to create a generic library and fix the compilation targets.

%\begin{figure}[ht]
%\centering
%\includegraphics[scale=0.325]{timeline_patch}
%\caption{Timeline of the PATCH submission process}
%\label{fig:timeline}
%\end{figure}

%The experience of submitting a patch was not as we expected. The process took 7 months and 16 days, significantly slower than expected. We initially submitted the patch on August 9, 2024, but we did not receive replies. Thus, we had to resend it on October 9, 2024. We received an initial response on November 3, 2024, requesting minimal changes. We submitted a second version with the requested changes on November 16, 2024. We did not receive replies initially and will need to resend the patch on February 7, 2025. We received a reply on February 11, 2025, asking for additional changes. We sent a third version on February 24, 2025, and the patch was accepted and integrated into the kernel codebase on February 25, 2025.
%Figure \ref{fig:timeline} illustrates the timeline of the patch submission.

%After the initial reply from the maintainers, we observed that the review process became more responsive. We hypothesize that once a patch receives initial attention, it is more closely tracked by maintainers. Another contributing factor to the initial delay could be the maintainers' workload at the time of the first submission.

%We discussed the Display driver with our point of contact at AMD to understand the reason for the slow patch submission process. We learned that

\section{Non-participant Observation Study}

@@

\subsection{Overview of the submitted patches}

@@

%\begin{figure}
%\centering
%\includegraphics[scale=0.8]{group_10_reply}
%\caption[Maintainers' Feedback on Group 10 Patch]{
% Maintainers' Feedback on Group Ten's Patch
%}
%\label{fig:group_10_reply}
%\end{figure}

@@

\subsection{Statistics of the contributions}

>>>>

\section{Methods and Tool}

We employ a multi-method research approach. We developed and evaluated a proposed tool using an approach from the literature, comparing it against the BigCloneBench dataset~\cite{bigclonebench}.
%
Given the validation of the tool capabilities, we proceed to understand the viability 
of mitigating the code duplication found by the tool. We conducted an ethnographic study 
using a participant-observation methodology to understand its viability.
%
In the first phase, the main author explored the process of mitigating a code duplication found in the AMD Display driver and sent it as a PATCH to interact with the Linux kernel community and get their feedback. In the second phase, we proposed to newcomers to mitigate the code duplications found by the tool in the IIO subsystem and the AMD Display driver and pass the process of sending the mitigations as a PATCH.

%Uniformizar este entendimento no texto:

%Fase 1:Complete Participant
%O pesquisador não estava observando outros desenvolvedores passarem pelo processo; ele se tornou o desenvolvedor. Ele se imergiu totalmente no papel de um contribuidor do kernel, enfrentando os mesmos desafios, normas sociais e barreiras técnicas que qualquer membro da comunidade enfrentaria.
% O objetivo aqui é ganhar uma visão "êmica" (de dentro), entendendo a cultura e o processo por meio da experiência vivida. Isso é, por vezes, também chamado de autoetnografia. Ele não estava "fingindo", ele estava genuinamente tentando contribuir para entender as dores.

%Fase 2: Participant-as-Observer
%O papel do pesquisador (no caso pesquisadores) mudou fundamentalmente. Ele deixou de ser o ator principal (o contribuidor) e passou a ser o facilitador (mentor).
%O foco dos dados coletados não é mais a autoexperiência do pesquisador, mas sim a experiência dos estudantes.
%A participação ainda é alta (selecionar, direcionar, revisar), o que influencia ativamente o cenário. No entanto, a identidade principal perante os participantes (os estudantes) é a de pesquisador/mentor, e o objetivo é observar o processo deles.

\subsection{Tool}
We developed \texttt{anonymous-tool}, a command-line tool designed to detect and analyze function-level code duplication within the Linux kernel to support our investigation. Released under the LGPL license, it employs a two-stage architecture that separates the computationally intensive task of finding duplicates from the process of querying the results. The first stage, the Preprocessor, analyzes the entire codebase, parses C functions, and stores identified duplicates in a database. The second stage, the Query Responder, uses this database to retrieve information about the identified duplications quickly.
%The tool is available at \textit{\href{https://github.com/arkanjo-tool/arkanjo}{github.com/arkanjo-tool/arkanjo}}.

\texttt{Anonymous-tool} is designed for the specific challenges of the kernel, with a focus on scalability to handle its size and prioritizing duplications that impact maintainability. The tool's primary role in this study was to provide a concrete entry point for new contributors by identifying function-level duplications to address. The tool uses a text similarity method that applies TF-IDF vector embedding, implemented with the Gensim library~\cite{gensim}, and computes cosine similarity to detect clones.

\subsection{Complete Participant Observation Study}
\label{sec:participant}

@@

@@

\subsection{Participant-as-Observer Study}

@@

\section{Results from Participant Observation}

%Results part 1 - AMD/Luan
%Ver como melhor iniciar.
The first duplicated function pair chosen is function \textit{offset\_to\_id} on files \textit{dc/gpio/dcn32/hw\_translate\_dcn32.c} and \textit{dc/gpio/dcn315/hw\_translate\_dcn315.c}. The second duplicated function pair is function \textit{phy\_id\_to\_atom} in file 
\textit{ \seqsplit{dc/bios/dce110/command\_table\_helper\_dce110.c}} and \textit{\seqsplit{dc/bios/dce60/command\_table\_helper\_dce60.c}}.
Hereafter, the mitigation on \textit{offset\_to\_id} will be referred to as mitigation 1, and the mitigation on 
\textit{phy\_id\_to\_atom} will be referred to as mitigation 2.

Firt, we found that the \textit{offset\_to\_id} function 
exists in 10 code files, enclosing multiple GPU architectures. The duplicated 
functions are not exactly equal, as some minimal specific logic is applied to each GPU architecture. To address this issue, we used the extract method to 
break the function into smaller parts and the parameterized method to address 
the specific logic for the GPUs.

Second, we found that the \textit{phy\_id\_to\_atom} function exists in five code files, and two other functions are duplicated 
in those five code files. All functions across the files were exactly equal, 
so there was no need to apply the refactoring methods presented in the literature. 
The refactoring was resumed to create a generic library and fix the compilation targets.

\section{Results from Non-Participant Observation}

@@

@@

%\subsection{Statistics of the contributions}
